### 2019-04-29 report/overview.tex
Advance in overview

@@

Since Sinkhorn can be parallelized, one can compute a large amount of Optimal
Transport distances: for instance look into a database for similarity based
retrieval, blablabla

@@

For image retrieval blablabla...

@@

For similar documents retrieval blabla

@@

Introduce Wasserstein distance into optimizations problem.

We show that we can computationaly differentiate, and therefore we can deal
with it in a variety of optimization problem, including machine learning
related opt. problems.

Somehow revisiting machine learning problems.

@@

histograms, distribution, images, experiments (repetition of same expriment,
MEG mesures, average over the experiments might be relevant)

@@

From a very big dataset, split it in $J$ parts, learn the distribution (eg. w/
Bayesaian Learning) from each of the part in parralel (on $J$ machines):
the $J$ distributions should not be very different, but still the variance in
the data will cause some variation: the idea is to use Wasserstein averaging
to aggregate the distributions, and get a more meaningful one. Theoretic
guaranties! :D

Wasserstein posterior

>>>>

Now that we have convinced ourselves of the good mathematical foundations
of Optimal Transport, and of the fact that we can actually compute it, we
can wonder how we can put this tool to great use in the setting of machine
learning.

@@

In the previous section we have seen that we have an algorithm to compute
optimal transport (Sinkhorn's algorithm), and that this algorithm can be
parallelized. We can take advantage of this fact to actually compute optimal
transport among huge databases. One of the application of this is similarity
based retrieval.

@@

The color-histogram of an image is a tool that characterizes the color
distribution of an image. Interestingly enough, it has been noticed that
color-histogram are good tool to describe the similarity of two images. Since
we now have at our disposal a tool to compute the distance between two
distribution probability, we can apply it for histograms (that are really not
that far from probability distribution, so much so that the algorithms and
methods that we described earlier are relevant even without any fondamental
change of setting). For a given image we can thus compute the distance between
its histogram and the histogram of images comming from a bigger database:
picking images that have the most similar histogram (in the sense of the
Wasserstein distance) will therefore allow some form of similarity based
image retrieval.

\begin{center}
	\includegraphics[scale=0.3]{image_histo.png}
    \captionof{figure}{Two images and their respective color-histograms,
    embedded in $\R^2$}
\end{center}

@@

We can apply the same technique as before, but instead of considering
color-histograms, we can try to find some kind of distribution coming from
other kind of documents, for example text documents. Some previous work,
like the $\texttt{word2vec}$ framework allow for embedding of natural
language words into an Euclidean space. A text therefore defines some sort
of distribution of probability over this embedded space. As for images we
can then compare this distribution to the distributions of other texts coming
from a database (still in the sense of Wasserstein distance), and hopefully
retrieve text dealing with similar topic, although there may be some
significant differences in the actual vocabulary of those texts. We take
advantage of the geometry of the distribution of probability in the embedded
text, that tries to ensure that similar distributions are coming from texts
dealing with similar topics.

@@

We can go further than just using the Wasserstein to compute distance between
distributions and use this as the main tool for the machine learning problems
that we considered (we previously emphasized retrieval tasks) by introducing
the Wasserstein distance on well studied machine learning settings and
problems. We thus shed a new light on those problems, and the results are
often fruitful. Some authors have coined the term $\textbf{Wassersteinization}$
for this process.

\DEF{
	$\textbf{Wassersteinization}$ is the process of introducing optimal
	transport into an optimization or machine learning problem
}

One way to approach this process is to introduce Wasserstein distance as a
loss or fidelity term in optimization problems. This will motivate the interest
toward the study of not only the very Wasserstein distance, but also of its
derivative. More precisely we can show that we can computationaly evaluate
the derivatives of the Wasserstein distance. Therefore we can deal with it a variety of optimization problems, including machine lerning related ones.

@@

We can also see optimal transport as a normalization tool for data. One of
the setting where this point of view as been proposed is the one of data
averaging. We have already discussed earlier the benefits of Wasserstein's
barycenters when compared to, say linear interpolations or $l_2$
barycenters. One can see this put in good use in the case of data acquired
through real world experiment. For instance, we can imagine a neurology
setting, where the goal is to study the reaction of the brain of a patient
to a certain precise stimulus. To do this we subjet the patient to the stimuli
several time and we aim to submit the measures of the brain activity to
a machine algorithm to study some of its features. But because of a whole
lot of data acquisition hazards (noise, imprecisions in the measures, etc),
the result will slighty differ from one experiment to another although it
concerns the same stimulus. Therefore computing the average of the brain
activity (that can easily seen to be analog to a probability distribution
on the brain space), in the sense of Wasserstein barycenters can hopefully
give access to more relevant data, without loosing the meaningful information.

\begin{center}
    \centering
	\includegraphics[scale=0.3]{brain_map.png}
    \captionof{figure}{On the left the average brain map among several
    experiments, on the right the brain activity map on a single experiment.}
\end{center}

The same idea can be followed in other fields, like imagery for instance.

@@

Instead of processing the input of our machine learning algorithms, one can
use this idea of aggregating several distribution on the output of machine
learning algorithms. Imagine that we want to use Bayesian learning (\emph{ie.}
learn a probability distribution, in a Bayesian framework), on a dataset so
huge that it cannot even fit on a single machine, and would it fit that the
computation time would be out of control. We can imagine to split the dataset
into $J$ different parts and distribute it amongst $J$ machines running the
same learning algorithms. The result of this process is a set of $J$
distributions learned from each part of the dataset. Although they represent
the same truth, because of the variation of the data those variation will
be slightly different. Once again Wasserstein-averaging provides a useful
tool to aggregate those distributions, into an hopefully meaningful one,
closer the underlying real distribution. The good news is that this Wasserstein
posterior aggregation method comes with theoretic statistical guaranties!
